---
title: "Homework4"
---

# Task 1

```{r}
library(tidyverse)

iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

2\. What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

Long format data means that per row there is only one observation with separate columns for the variables. Wide format data normally has multiple measurements of the same observation that is spread across different columns. Long format is usually preferable for statistical analysis because functions work well in R with this and it also makes it easier to manipulate, summarize, and analyze data.

# Task 2

## Glass Data

```{r}
library(tidyverse)

glass <- read_csv(
  "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = FALSE
)

names(glass) <- c(
  "Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe",
  "Type_of_glass"
)

glass
```

```{r}
glass <- glass |>
  mutate(
    Type_of_glass = factor(
      Type_of_glass,
      levels = c(1, 2, 3, 4, 5, 6, 7),
      labels = c(
        "building_windows_float_processed",
        "building_windows_non_float_processed",
        "vehicle_windows_float_processed",
        "vehicle_windows_non_float_processed",
        "containers",
        "tableware",
        "headlamps"
      )
    )
  ) |>
  filter(
    Fe < 0.2,
    Type_of_glass %in% c("tableware", "headlamps")
  )

glass
```

## Yeast Data

```{r}
yeast <- read_table(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
  col_names = c(
    "seq_name", "mcg", "gvh", "alm", "mit",
    "erl", "pox", "vac", "nuc", "class"
  )
)

yeast
```

```{r}
yeast <- yeast |>
  select(-seq_name, -nuc)

yeast
```

```{r}
yeast <- yeast |>
  group_by(class) |>
  mutate(
    across(
      c(mcg, gvh, alm, mit, erl, pox, vac),
      list(mean = mean, median = median),
      na.rm = TRUE
    )
  )

yeast
```

## Wine Data

```{r}
library(readxl)

white_wine <- read_excel("white-wine.xlsx") |>
  mutate(type = "white")

red_wine <- read_delim(
  "red-wine.csv",
  delim = ";",
  locale = locale(decimal_mark = ".")
) |>
  mutate(type = "red")

wine <- bind_rows(red_wine, white_wine)

wine
```

```{r}
wine <- wine |>
  filter(quality > 6.5, alcohol < 13.2)

wine
```

```{r}
wine <- wine |>
  arrange(desc(quality))

wine
```

```{r}
wine <- wine |>
  select(contains("acid"), alcohol, type, quality) |>
  group_by(quality) |>
  mutate(
    alcohol_mean = mean(alcohol),
    alcohol_sd = sd(alcohol)
  )

wine
```

# Task 3

```{r}
library(DBI)

con <- dbConnect(
  RSQLite::SQLite(),
  "PiyushPatil/Homework4/data/lahman.db"
)

dbListTables(con)
```

```{r}
teams_2015 <- tbl(con, "Teams") |>
  filter(yearID == 2015)

teams_2015
```

```{r}
teams_2015_sql <- tbl(
  con,
  sql("SELECT * FROM Teams WHERE yearID = 2015")
)

teams_2015_sql
```

```{r}
hall_of_fame <- tbl(con, "HallOfFame") |>
  select(playerID, yearID, category, inducted)

hall_of_fame
```

```{r}
hall_of_fame_people <- hall_of_fame |>
  left_join(
    tbl(con, "People") |>
      select(playerID, nameFirst, nameLast),
    by = "playerID"
  )

hall_of_fame_people
```

```{r}
manager_stats <- tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(
    Total_G = sum(G),
    Total_W = sum(W),
    Total_L = sum(L)
  ) |>
  mutate(
    Total_Win_Percentage = Total_W * 1.0 / Total_G
  ) |>
  arrange(desc(Total_Win_Percentage))

manager_stats
```

```{r}
hall_of_fame_managers <- hall_of_fame_people |>
  filter(inducted == "Y") |>
  inner_join(
    manager_stats,
    by = "playerID"
  )

hall_of_fame_managers
```
