---
title: "Homework 2 - Grace Haley"
format: html
---

## Question 1

Create vectors for the pre-treatment and post-treatment blood pressure measurements.

```{r}
pre_bp <- c(
  130, 128, 116, 124, 133,
  134, 118, 126, 114, 127,
  141, 138, 128, 140, 137,
  131, 120, 128, 139, 135
)

post_bp <- c(
  114, 98, 113, 99, 107,
  116, 113, 111, 119, 117,
  101, 119, 130, 122, 106,
  106, 124, 102, 117, 113
)
```

## Question 2

Assign subject names to each vector.

```{r}
patient_names <- paste("Subject", 1:20, sep = "_")

names(pre_bp) <- patient_names
names(post_bp) <- patient_names

pre_bp
post_bp
```
## Question 3

Calculate the change in blood pressure for each patient by subtracting the post-treatment measurements from the pre-treatment measurements.

```{r}
diff_bp <- pre_bp - post_bp
diff_bp
```

## Question 4

Calculate the average decrease in blood pressure across all patients.

```{r}
mean(diff_bp)
```

## Question 5

Determine which patients experienced a decrease in blood pressure after treatment using the `which()` function.

```{r}
which(diff_bp > 0)
```

## Question 6

Subset the vector of differences to return only those with a positive change.

```{r}
positive_changes <- diff_bp[diff_bp > 0]

positive_changes
```

## Question 7

Calculate the average decrease in blood pressure for only those patients whose blood pressure decreased.

```{r}
mean(positive_changes)
```
---
