---
title: "Homework 3"
author: "AMJ"
---

### Task 1

1.  I would use this path `../MyData.csv`.
2.  The major benefits of using R projects include:

- R projects keeps all my different project file types in one folder.
- R projects is integrated with tools like Git so I can easily use the terminal to upload and share my work.
- R projects provides a nice UI to keep up with all my project files and data in one place.

3.  Git is a version control tool that allows you to track changes in your code on your computer. Github, on the other hand, is a website that allows you to upload these changes to share with others. You can also download (pull) other Github users' code.
4.  A tibble doesn't show the entire dataset like a dataframe would when you print it, and it rather shows only rows that fit on the screen. When you access one column of a tibble, it will keep the tibble structure, but when you access one column of a dataframe, it will become a vector.

### Task 2

This code section is for the glass data.

1.  

```{r}
# I am importing the tidyverse library
library(tidyverse)
# This code creates a tibble called glass from the specified URL for the glass data.
# I assigned the column names just from the homework document. The delimiter is a comma.
# Read_csv is used since the data is separated by commas. It's an excel style file.
glass <- read_csv(
  "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c(
    "ID", "RI", "Na", "Mg", "Al", "Si",
    "K", "Ca", "Ba", "Fe", "TypeGlass"
  )
)
# This just prints the tibble object I just created called glass.
glass
```

The next code section is for the yeast data.

2.  

```{r}
# No need to import tidyverse again...
# This code creates a tibble called yeast from the specified URL for the yeast data.
# I assigned column names once again and I used read_table for the new space delimiter.
yeast <- read_table(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
  col_names = c(
    "Sequence_Name", "mcg", "gvh", "alm",
    "mit", "erl", "pox", "vac", "nuc", "Class"
  )
)
# Printing the tibble object yeast...
yeast
```

### Task 3

3.  

```{r}
# I first downloaded the white-wine.xlsx file and placed it in this folder.
# Importing readxl and dplyr library.
library(readxl)
library(dplyr)
# Since the data is in an excel format delimited by commas, we use read_excel.
white_wine <- read_excel("white-wine.xlsx", sheet = 1)
# I am calling the white_wine tibble I just made.
white_wine
```

4.  Next, I'm reading the second excel sheet and changing the column names to the ones from alternate column names from the second sheet.

```{r}
# I am reading the columns from the second sheet.
wine_names <- read_excel("white-wine.xlsx", sheet = 2)
# This replaces the column names in our tibble with the new wine_names column names.
colnames(white_wine) <- wine_names[[1]]
# (re)printing our tibble with the updated names. Now, the column names are a bit different.
white_wine

```

5.  Next, I'm adding a column to indicate that these are all white wines so far.

```{r}
white_wine <- white_wine %>%
  mutate(type = "white")
# Printing object to see if it worked.
white_wine
```

6.  Next, I will be adding the red wine data.

```{r}
# Reading the red_wine file which uses semicolon as a delimiter for data.
red_wine <- read_delim(
  "https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
  delim = ";"
)
# Printing object...
red_wine
# This indicates that this data is for red wine in a new column (like we just did for white).
colnames(red_wine) <- wine_names[[1]]
red_wine <- red_wine %>%
  mutate(type = "red")
# Printing object...
red_wine
```

7.  Finally, I will combine the two data sets into one data set called "wine".

```{r}
# This is done using the dplyr::bind_rows() function.
wine <-  bind_rows(white_wine, red_wine)
# Printing final object...
wine

```

### Task 4

8.  This is my IQR function which which you can pass unnamed arguments to. IQR is 75th percentile minus 25th percentile.

```{r}
getIQR <- function(x, ...) {
  q <- quantile(x, probs = c(0.25, 0.75), ...)
  # This line removes the labels on the returned vector result.
  unname(q[2] - q[1])
}
```

9.  This code runs the getIQR function twice on the specified columns from my wine object and displays the results.

```{r}
getIQR(wine$fixed_acidity)
getIQR(wine$citric_acid)
```

10. This is my getCoefVar function which you can pass unnamed arguments to. Coefficient of variation is standard deviation divided by the absolute value of mean.

```{r}
getCoefVar <- function(x, ...) {
  sd(x, ...) / abs(mean(x, ...))
}
```

11. This code runs the getCoefVar function twice on the specified columns from my wine object and displays the results.

```{r}
getCoefVar(wine$fixed_acidity)
getCoefVar(wine$citric_acid)
```

12. Here, I'm creating a function called getSummaries which calls both the IQR and CoefVar functions. There's no need to rewrite them since we can just call them within this function.

```{r}
getSummaries <- function(x, ...) {
  # Have to make sure to label them to not get confused.
  c(
    IQR = getIQR(x, ...),
    CoefVar = getCoefVar(x, ...)
  )
}
```

13. This code runs the getSummaries function twice on the specified columns from my wine object and displays the results.

```{r}
getSummaries(wine$fixed_acidity)
getSummaries(wine$citric_acid)
```

14. Finally, I am creating my own infix function which will be used instead of cbind. Cbind combines data from vectors, data frames, etc., via columns (so horizontally).

```{r}
# This is my new cbind infix function which takes two parameteres x and y.
`%cbind%` <- function(x, y) {
  cbind(x, y)
}
# This is just the code given in the homework page for us to test the new function...
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)
# Testing infix function with dataframe1 and dataframe2...
df1 %cbind% df2
# Testing infix function with dataframe1, dataframe2, and vector1...
df1 %cbind% df2 %cbind% vec1
```
