---
title: "Homework 5 Combined"
---

# Task 1 - Querying an API and a Tidy-Style Function 

## Question 1

```{r}
library(httr)

api_key <- Sys.getenv("NEWS_API_KEY")

# Send a GET request to the NewsAPI for articles about AI 

news <- GET(
  paste0(
    "https://newsapi.org/v2/everything?q=artificial%20intelligence",
    "&from=2026-09-01",
    "&language=en",
    "&pageSize=50",
    "&apiKey=", api_key
  )
)

news
```

## Question 2

```{r}
# Extracting the article info from the API reponse

news_content <- content(news, simplifyDataFrame = TRUE)

news_content$articles

str(news_content$articles)
```

## Question 3

```{r}
# This function searches NewsAPI using a subject and starting date 

get_news <- function(subject, from_date, api_key) {
  
  subject <- URLencode(subject)
  
  url <- paste0(
    "https://newsapi.org/v2/everything?q=", subject,
    "&from=", from_date,
    "&language=en&pageSize=50&apiKey=", api_key
  )
  
  news <- GET(url)
  
  news_content <- content(news, simplifyDataFrame = TRUE)
  
  news_content$articles
}

get_news("artificial intelligence", "2026-09-01", api_key)

get_news("Apple", "2026-09-01", api_key)
```

# Task 2 - Summarizing Roller Coaster Data

```{r}
library(tidyverse)

coasters <- read_csv(
  "https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv"
)

# Convert categorical variables to factors
coasters <- coasters %>%
  mutate(
    `Census Region` = as.factor(`Census Region`),
    AgeGroup = as.factor(AgeGroup),
    Scale = as.factor(Scale),
    Type = as.factor(Type),
    Design = as.factor(Design),
    `Inversions?` = as.factor(`Inversions?`),
    Make = as.factor(Make)
  )
```

## Question 4

```{r}
# Checks the structure
str(coasters)
```

The data contains 635 observations and 21 variables. The categorical variables used in the analysis are stored as factors, while the quantitative variables are stored as numeric variables. The variables appear to be stored appropriately for the analysis.

## Question 5

```{r}
# Counts the number of missing values in each variable
colSums(is.na(coasters))
```

The variables with missing values are TopSpeed with 74 missing values, MaxHeight with 8, Drop with 128, Length with 33, Duration with 66, and Make with 36. All other variables have no missing values.

## Question 6

```{r}
# One-way contingency table
table(coasters$Type)

# Two-way contingency table
table(coasters$Type, coasters$Scale)

# Three-way contingency table
table(coasters$Type, coasters$Scale, coasters$`Census Region`)
```

For the one-way table, the value 530 means that there are 530 steel roller coasters in the dataset.

For the two-way table, the value 216 means that there are 216 steel roller coasters classified as Extreme.

For the three-way table, the value 42 means that there are 42 steel roller coasters classified as Extreme in the Midwest.

## Question 7

```{r}
# Method 1: Filter the data first
south_coasters <- coasters %>%
  filter(`Census Region` == "South")

table(south_coasters$Type, south_coasters$Scale)

# Method 2: Create a three-way table and subset it
coaster_table <- table(
  coasters$Type,
  coasters$Scale,
  coasters$`Census Region`
)

coaster_table[, , "South"]
```

Both methods produce the same conditional two way table for roller coasters in the south

## Question 8

```{r}
coasters %>%
  group_by(Type, Scale) %>%
  summarize(Count = n(), .groups = "drop") %>%
  pivot_wider(
    names_from = Scale,
    values_from = Count
  )
```

## Question 9

```{r}
# Stacked bar graph
ggplot(coasters, aes(x = Type, fill = Scale)) +
  geom_bar() +
  labs(
    title = "Roller Coaster Scale by Type",
    x = "Type",
    y = "Count",
    fill = "Scale"
  )

# Side-by-side bar graph
ggplot(coasters, aes(x = Type, fill = Scale)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Roller Coaster Scale by Type",
    x = "Type",
    y = "Count",
    fill = "Scale"
  )
```
