---
title: "Homework 3 Combined"
author: "Group 9"
---

# Task 1: Conceptual Questions

1.  If the working directory is `myfolder/homework/`, the relative path needed to access `myfolder/MyData.csv` is `../MyData.csv`. The `..` moves up one directory from the `homework` folder to the `myfolder` folder.

2.  Some major benefits of using R Projects are that they keep all files for a project organized in one location, which make relative file paths easier to use and provide a consistent working directory. This also makes it easier to use Git and GitHub for a project.

3.  Git is a version control system that tracks changes made to files and allows users to keep different versions of their work. GitHub is an online platform that hosts Git repositories and makes it easy to store, share, and collaborate on projects.

4.  Two main differences between a tibble and a data.frame are:

    - Tibbles print only the rows and columns that fit on the screen instead of the entire data set.
    - Tibbles are more strict about how they handle variables and do not automatically change things such as variable names or data types in the way traditional data frames might.

# Task 2: Reading Delimited Data

## Question 1: Glass Data

```{r}
# I am importing the tidyverse library
library(tidyverse)
# This code creates a tibble called glass from the specified URL for the glass data.
# I assigned the supplied column names. The delimiter is a comma.
# Read_csv is used since the data is separated by commas. It's an excel style file.
glass <- read_csv(
  "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c(
    "ID", "RI", "Na", "Mg", "Al", "Si",
    "K", "Ca", "Ba", "Fe", "TypeGlass"
  )
)
# This just prints the tibble object I just created called glass.
glass
```

## Question 2: Yeast Data

```{r}
# No need to import tidyverse again...
# This code creates a tibble called yeast from the specified URL for the yeast data.
# I assigned the supplied column names once again and I used read_table for the new space delimiter.
yeast <- read_table(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
  col_names = c(
    "Sequence_Name", "mcg", "gvh", "alm",
    "mit", "erl", "pox", "vac", "nuc", "Class"
  )
)
# Printing the tibble object yeast...
yeast
```

# Task 3: Combining Excel and Delimited Data

## Question 3

I imported the white wine data from the first sheet of the Excel file using the readxl package and printed the resulting tibble.

```{r}
library(readxl)

white_wine <- read_excel("white-wine.xlsx", sheet = 1)

white_wine
```

## Question 4

I read the alternative variable names from the second Excel sheet and used them to replace the original column names in the white wine data.

```{r}
wine_names <- read_excel("white-wine.xlsx", sheet = 2)

colnames(white_wine) <- wine_names[[1]]

white_wine
```

## Question 5

I added a type column to identify every observation in the white wine data as "white".

```{r}
white_wine <- white_wine |>
  mutate(type = "white")

white_wine
```

## Question 6

I read the semicolon-delimited red wine data directly from the URL. I then replaced its column names using the same names as the white wine data and added a type column identifying the observations as "red".

```{r}
red_wine <- read_delim(
  "https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
  delim = ";"
)

colnames(red_wine) <- wine_names[[1]]

red_wine <- red_wine |>
  mutate(type = "red")

red_wine
```

## Question 7

Since the red and white wine data now have the same columns, I combined them into one tibble using bind_rows().

```{r}
wine_data <- bind_rows(white_wine, red_wine)

wine_data
```

# Task 4: Writing a Function

## Question 8

I created the getIQR() function to calculate the interquartile range by subtracting the 25th percentile from the 75th percentile. I used unname() so the returned value does not have a percentile name.

```{r}
getIQR <- function(x, ...) {
  q <- quantile(x, probs = c(0.25, 0.75), ...)
  # This line removes the labels on the returned vector result.
  unname(q[2] - q[1])
}
```

## Question 9

I applied getIQR() to the fixed_acidity and citric_acid columns of the combined wine data.

```{r}
getIQR(wine_data$fixed_acidity)
getIQR(wine_data$citric_acid)
```

## Question 10

I created getCoefVar() to calculate the sample coefficient of variation by dividing the sample standard deviation by the absolute value of the sample mean.

```{r}
getCoefVar <- function(x, ...) {
  sd(x, ...) / abs(mean(x, ...))
}
```

## Question 11

I applied getCoefVar() to the same two variables used in Question 9.

```{r}
getCoefVar(wine_data$fixed_acidity)
getCoefVar(wine_data$citric_acid)
```

## Question 12

I created the wrapper function getSummaries() to return both the IQR and coefficient of variation. Instead of repeating the calculations, the function calls the two functions created above.

```{r}
getSummaries <- function(x, ...) {
  # Have to make sure to label them to not get confused.
  c(
    IQR = getIQR(x, ...),
    CoefVar = getCoefVar(x, ...)
  )
}
```

## Question 13

I applied getSummaries() to fixed_acidity and citric_acid to return both summary measures for each variable.

```{r}
getSummaries(wine_data$fixed_acidity)
getSummaries(wine_data$citric_acid)
```

## Question 14

I created the infix function %c% as an alternative way to call cbind(). I first tested it on the two data frames and then tested it on both data frames and the vector.

```{r}
# This is my new cbind infix function which takes two parameteres x and y.
`%cbind%` <- function(x, y) {
  cbind(x, y)
}
# This is just the code given in the homework page for us to test the new function...
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)
# Testing infix function with dataframe1 and dataframe2...
df1 %cbind% df2
# Testing infix function with dataframe1, dataframe2, and vector1...
df1 %cbind% df2 %cbind% vec1
```
