---
title: "Homework 5"
author: "AMJ"
---

# Task 1: Querying an API and a Tidy-Style Function

1.  I'm using my API key to get information about my topic. Calling articles_df returns a tibble with the information I'm looking for like title, author, etc.

```{r}
library(httr)
library(dplyr)

key = '617117eac1654fa2a323dd9021ca4f71'


news_response <- GET(
  "https://newsapi.org/v2/everything",
  query = list(
    q = "Iran",
    from = "2026-09-01",
    language = "en",
    pageSize = 50,
    apiKey = key
  )
)

news_content <- content(news_response)
articles_df <- bind_rows(news_content$articles)
#2. Below is the data frame that is generated.
articles_df


```

3.  This function can be called with the 3 parameters to get information about a certain topic.

```{r}
get_news <- function(topic, start_date, api_key) {
  
  response <- GET(
    "https://newsapi.org/v2/everything",
    query = list(
      q = topic,
      from = start_date,
      language = "en",
      pageSize = 50,
      apiKey = key
    )
  )
  
  news_content <- content(response)
  articles_df <- bind_rows(news_content$articles)
  return(articles_df)
}
```

This object is a function call of the function I just made and it results in news about cancer.

```{r}
cancer_news <- get_news(
  "cancer",
  "2026-09-01",
  key
)
cancer_news
```

This object is a function call of the function I just made and it results in news about Kazakhstan.

```{r}

kazakhstan_news <- get_news(
  "kazakhstan",
  "2026-09-01",
  key
)
kazakhstan_news
```

# Task 2: Summarizing Roller Coaster Data

```{r}
# The csv file is being read and factors were created.
library(tidyverse)
coasters <- read_csv("635USrollercoasters.csv")
coasters <- coasters |>
  mutate(
    Park = factor(Park),
    City = factor(City),
    State = factor(State),
    `Census Region` = factor(`Census Region`),
    AgeGroup = factor(AgeGroup),
    Scale = factor(Scale),
    Type = factor(Type),
    Design = factor(Design),
    `Inversions?` = factor(`Inversions?`),
    Make = factor(Make)
  )
```

```{r}
# 4. I'm looking into the data here. There are 635 rows and 21 columns of data. Everything seems to be stored properly and the data makes sense.
str(coasters)
glimpse(coasters)
```

5.  This code returns the variables with missing columns and how many missing values there are.

```{r}
coasters |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(
    everything(),
    names_to = "Variable",
    values_to = "Missing"
  ) |>
  filter(Missing > 0)
```

There are missing values in TopSpeed, MaxHeight, Drop, Length, Duration, and Make. The missing column shows how many missing values there are. I went ahead and looked at the Excel sheet to make sure and counted up the missing records in TopSpeed and there was exactly 74. There are a lot of blank rows in the csv file but the code skips those and doesn't count them toward the missing values.

6. Here, I am creating a one-way, two-way, and three-way contingency tables using table().

```{r}
table(coasters$Type)
```
The first value, 530, means that there are 530 steel roller coasters in the dataset.
```{r}
table(coasters$Type, coasters$Scale)
```
The first value, 216, means that there are 216 steel roller coasters that are classified as "extreme" in the dataset.
```{r}
table(coasters$Type, coasters$Scale, coasters$AgeGroup)
```
The first value, 8, means that there are 8 steel roller coasters that are classified as "extreme" and are classified as "older" in the dataset.

7. I am creating a conditional two-way table using two methods.

Here, I am creating a two-way contingency table by subsetting the data using filter() and then creating the two-way table.
```{r}
older_coasters <- coasters |>
  filter(AgeGroup == "1:Older")
table(older_coasters$Type, older_coasters$Scale)
```
Here, I am creating a three-way table first and then subsetting it.
```{r}
three_way <- table(
  coasters$Type,
  coasters$Scale,
  coasters$AgeGroup
)
three_way[, , "1:Older"]
```
As you can see, both resulting tables match.

8. Here, I am creating a two-way contingency table using group_by() and summarize() from dplyr and using pivot_wider to change the try to match the resulting tibble to a table.

```{r}
two_way_dplyr <- coasters |>
  group_by(Type, Scale) |>
  summarize(Count = n(), .groups = "drop") |>
  pivot_wider(
    names_from = Scale,
    values_from = Count
  )

two_way_dplyr
```

9. I am creating a stacked bar graph and a side-by-side bar graph with labels.

This is the stacked bar graph.
```{r}
ggplot(coasters, aes(x = Type, fill = Scale)) +
  geom_bar() +
  labs(
    x = "Roller Coaster Type",
    y = "Number of Roller Coasters",
    title = "Roller Coaster Scale by Type"
  )
```


This is the side-by-side bar graph.
```{r}
ggplot(coasters, aes(x = Type, fill = Scale)) +
  geom_bar(position = "dodge") +
  labs(
    x = "Roller Coaster Type",
    y = "Number of Roller Coasters",
    title = "Roller Coaster Scale by Type"
  )
```
