---
title: "Homework 4"
author: "Group 9"
---

# Task 1: Conceptual Questions

1.  Here, we're using the Base R pipe \|\> to take the result from the previous line and put it into the next function over and over. This is called a chain.

```{r}
library(dplyr)
iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```



2.  

- Long format data is when repeated measurements are placed in multiple rows with one column identifying the measurement and another containing its value.
- Wide format data is when one observational unit is represented by one row, with repeated measurements spread across multiple columns.
- For statistical analysis, we generally prefer long format data.

# Task 2: Glass, Yeast, and Wine Data

I am just using the code from HW3 to read the glass data.
```{r}
library(tidyverse)
# This code creates a tibble called glass from the specified URL for the glass data.
# I assigned the supplied column names. The delimiter is a comma.
# Read_csv is used since the data is separated by commas. It's an excel style file.
glass <- read_csv(
  "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c(
    "ID", "RI", "Na", "Mg", "Al", "Si",
    "K", "Ca", "Ba", "Fe", "TypeGlass"
  )
)
```

3. Here, I am starting a chain that will overwrite the TypeGlass variable using mutate.
4. At the end, there is code to filter observations so that only ones where Fe<0.2 and TypeGlass is either 'tableware' or 'headlamp'.

```{r}
# Done using Dplyr chain.
glass <- glass |>
  mutate(
    TypeGlass = factor(
      TypeGlass,
      levels = c(1, 2, 3, 4, 5, 6, 7),
      labels = c(
        "building_windows_float_processed",
        "building_windows_non_float_processed",
        "vehicle_windows_float_processed",
        "vehicle_windows_non_float_processed",
        "containers",
        "tableware",
        "headlamps"
      )
    )
  ) |>
  filter(
    Fe < 0.2,
    TypeGlass %in% c("tableware", "headlamps")
  )
glass
```


  
This is code that just reads the yeast data.
```{r}
yeast <- read_table(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
  col_names = c(
    "seq_name", "mcg", "gvh", "alm",
    "mit", "erl", "pox", "vac", "nuc", "class"
  )
)
```

5. I am starting a chain that removes seq_name and nuc columns.  
6. The chain continues and I added the specified columns. There is mean and median for each numeric variable and a class grouping.

```{r}
# Done again using Dplyr chain.
#5.
yeast |>
  select(-seq_name, -nuc) |>
  group_by(class) |>
  # 6. Using mutate now..
  mutate(
    across(
      c(mcg, gvh, alm, mit, erl, pox, vac),
      list(mean = mean, median = median)
    )
  )
```

This is just reading the wine data and combining it from HW3. It's the same code.

```{r}
library(readxl)

white_wine <- read_excel("white-wine.xlsx", sheet = 1)
wine_names <- read_excel("white-wine.xlsx", sheet = 2)

colnames(white_wine) <- wine_names[[1]]
white_wine <- white_wine |>
  mutate(type = "white")
red_wine <- read_delim(
  "https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
  delim = ";"
)

colnames(red_wine) <- wine_names[[1]]

red_wine <- red_wine |>
  mutate(type = "red")
wine_data <- bind_rows(white_wine, red_wine)
```

7. I started a chain to look at wine records with quality > 6.5 and alcohol > 13.2   

8. Now, the quality is arranged from highest to lowest (descending).

9. Next, it selects from where the variables contains 'acid'. And also the alcohol, type, and quality variables.

10. The chain continues with adding the mean and standard deviation of the alcohol variable to the data set of the quality variable.

```{r}
# Starting Dplyr chain..
wine_data |>
  #7.
  filter(quality > 6.5, alcohol < 13.2) |>
  #8.
  arrange(desc(quality)) |>
  #9.
  select(contains("acid"), alcohol, type, quality) |>
  #10. Using mutate now..
  group_by(quality) |>
  mutate(
    mean_alcohol = mean(alcohol),
    sd_alcohol = sd(alcohol)
  )
```

# Task 3: Database Practice

11. I am connecting to the database and displaying the tables.

```{r}
# Code is from the online textbook.
library(DBI)
con <- dbConnect(RSQLite::SQLite(), "lahman.db")
dbListTables(con)
```

12. I used the tbl function to return all 2015 data from the Teams table. This was done with Dplyr.

```{r}
teams_2015 <- tbl(con, "Teams") |>
  filter(yearID == 2015) |>
  collect()

teams_2015
```

13. I used the same tbl function but used SQL instead here.

```{r}
teams_2015_sql <- tbl(
  con,
  sql("SELECT * FROM Teams WHERE yearID = 2015")
) |>
  collect()

teams_2015_sql
```

14. This code returns hall of fame data including playerID, yearID, and category.

```{r}
hof <- tbl(
  con,
  # SQL statement
  sql("
    SELECT playerID, yearID, category
    FROM HallOfFame
    WHERE inducted = 'Y'
  ")
) |>
  collect()

hof
```

15. This code combines the previous table with the People table so we can see the player names.

```{r}
hof_names <- tbl(
  con,
  # SQL Statement
  sql("
    SELECT H.playerID, H.yearID, H.category,
           P.nameFirst, P.nameLast
    FROM HallOfFame AS H
    LEFT JOIN People AS P
      ON H.playerID = P.playerID
    WHERE H.inducted = 'Y'
  ")
) |>
  collect()

hof_names
```

16. This code returns the playerID, and G, W, and L columns from the managers table. The chain also calculates overall win/loss records for the same records.

```{r}
manager_records <- tbl(
  con,
  # SQL Statement
  sql("
    SELECT playerID,
           SUM(G) AS G_managed,
           SUM(W) AS Total_W,
           SUM(L) AS Total_L
    FROM Managers
    GROUP BY playerID
  ")
) |>
  # Using Dplyr to calculate their win loss and arrange it.
  collect() |>
  mutate(
    Win_Loss_Percentage = Total_W / G_managed
  ) |>
  arrange(desc(Win_Loss_Percentage))

manager_records
```

17. This code returns the people that managed a team and got a HOF induction. The category doesn't have to be Manager though.

```{r}
# Just reusing the objects created above! No need to rewrite everything.
hof_managers <- inner_join(
  hof_names,
  manager_records,
  by = "playerID"
)

hof_managers
```
