---
title: "Homework 2 (Combined)"
format: html
---

# Task 1


## Question 1

Create vectors for the pre-treatment and post-treatment blood pressure measurements.

```{r}
pre_bp <- c(
  130, 128, 116, 124, 133,
  134, 118, 126, 114, 127,
  141, 138, 128, 140, 137,
  131, 120, 128, 139, 135
)

post_bp <- c(
  114, 98, 113, 99, 107,
  116, 113, 111, 119, 117,
  101, 119, 130, 122, 106,
  106, 124, 102, 117, 113
)
```

## Question 2

Assign subject names to each vector.

```{r}
patient_names <- paste("Subject", 1:20, sep = "_")

names(pre_bp) <- patient_names
names(post_bp) <- patient_names

pre_bp
post_bp
```

## Question 3

Calculate the change in blood pressure for each patient by subtracting the post-treatment measurements from the pre-treatment measurements.

```{r}
diff_bp <- pre_bp - post_bp
diff_bp
```

## Question 4

Calculate the average decrease in blood pressure across all patients.

```{r}
mean(diff_bp)
```

## Question 5

Determine which patients experienced a decrease in blood pressure after treatment using the `which()` function.

```{r}
which(diff_bp > 0)
```

## Question 6

Subset the vector of differences to return only those with a positive change.

```{r}
positive_changes <- diff_bp[diff_bp > 0]
positive_changes
```

## Question 7

Calculate the average decrease in blood pressure for only those patients whose blood pressure decreased.

```{r}
mean(positive_changes)
```

# Task 2


This task combines `patient`, `pre_bp`, `post_bp`, and `diff_bp` into a single data frame, pulls out the patients whose blood pressure difference is negative, and adds a logical column flagging whether `post_bp` is under 120.

```{r}
bp_df <- data.frame(
  patient = patient_names,
  pre_bp = as.numeric(pre_bp),
  post_bp = as.numeric(post_bp),
  diff_bp = as.numeric(diff_bp),
  stringsAsFactors = FALSE
)
rownames(bp_df) <- NULL

# patients whose blood pressure difference is negative
bp_df[bp_df$diff_bp < 0, ]

bp_df$post_under_120 <- bp_df$post_bp < 120
bp_df

knitr::kable(bp_df)
```

# Task 3



The placebo group vectors are created and combined into a data frame with the same structure as `bp_df`. Both data frames are then stored together in a list.

```{r}
placebo_pre <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
placebo_post <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
placebo_names <- paste("Subject", 1:10, sep = "_")

placebo_df <- data.frame(
  patient = placebo_names,
  pre_bp = placebo_pre,
  post_bp = placebo_post
)

placebo_df$diff_bp <- placebo_df$pre_bp - placebo_df$post_bp
placebo_df$post_under_120 <- placebo_df$post_bp < 120

placebo_df
```

```{r}
bp_list <- list(
  treatment = bp_df,
  placebo = placebo_df
)
```

Accessing the treatment data frame three different ways.

```{r}
bp_list[1]
bp_list[[1]]
bp_list$treatment
```

Accessing the pre-treatment blood pressure column from the placebo data frame.

```{r}
bp_list$placebo$pre_bp
```

# Task 4


A `status` column is added to each data frame in the list and filled in with a for loop: `"optimal"` when `post_bp <= 120`, `"borderline"` when `120 < post_bp <= 130`, and `"high"` when `post_bp > 130`.

```{r}
bp_list$treatment$status <- character(nrow(bp_list$treatment))
bp_list$placebo$status <- character(nrow(bp_list$placebo))

for (i in 1:nrow(bp_list$treatment)) {
  if (bp_list$treatment$post_bp[i] <= 120) {
    bp_list$treatment$status[i] <- "optimal"
  } else if (bp_list$treatment$post_bp[i] <= 130) {
    bp_list$treatment$status[i] <- "borderline"
  } else {
    bp_list$treatment$status[i] <- "high"
  }
}

for (i in 1:nrow(bp_list$placebo)) {
  if (bp_list$placebo$post_bp[i] <= 120) {
    bp_list$placebo$status[i] <- "optimal"
  } else if (bp_list$placebo$post_bp[i] <= 130) {
    bp_list$placebo$status[i] <- "borderline"
  } else {
    bp_list$placebo$status[i] <- "high"
  }
}

bp_list$treatment
bp_list$placebo
```
