---
title: "Task1"
---

# Task 1

Vector creation for both pre and post data

```{r}
pre_data <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

post_data <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)


```

Assigning names for both pre and post data

```{r}
names(pre_data) <- paste("Subject", 1:20, sep="_")
names(post_data) <- paste("Subject", 1:20, sep="_")
```

Creating a vector that holds the difference between pre and post data

```{r}
diff_bp <- pre_data - post_data
```

Finding an average of differences between pre and post data

```{r}
avg <- mean(diff_bp)
```

Which function to determine which of the differences are positive

```{r}
which(diff_bp > 0)
```

New Vector that stores only positive differences

```{r}
positive_bp <- diff_bp[which(diff_bp > 0)]
```

Finding average of only the positive differences

```{r}
positive_avg <- mean(positive_bp)
```

# Task 2

Creating a data frame to hold all the data

```{r}
bp_df <- data.frame(
  patient = names(pre_data),
  pre_bp = pre_data,
  post_bp = post_data,
  diff_bp = diff_bp
)
```

Creating a vector that only holds negative difference values

```{r}
subset(bp_df, diff_bp < 0)
```

Column that holds only values under 120 for post data is added

```{r}
bp_df$post_bp_under120 <- bp_df$post_bp < 120
bp_df
```

```{r}
knitr::kable(bp_df)
```

# Task 3

New data for placebo pre and post created along with a difference vector

```{r}
placebo_pre <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
placebo_post <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
placebo_diff <- placebo_pre - placebo_post

```

Assigning names

```{r}
names(placebo_pre) <- paste("Subject", 1:10, sep = "_")
names(placebo_post) <- paste("Subject", 1:10, sep = "_")

```

Data frame similar that holds all the data

```{r}
placebo_df <- data.frame(
  patient = names(placebo_pre),
  pre_bp = placebo_pre,
  post_bp = placebo_post,
  diff_bp = placebo_diff
)
```

A column is added to the data frame that depicts if post data is under 120

```{r}
placebo_df$post_bp_under120 <- placebo_df$post_bp < 120
placebo_df
```

Creation of a list that holds previous and current data

```{r}
bp_list <- list(
  treatment = bp_df,
  placebo = placebo_df
)
```

Accessing the first element of the list in three ways

```{r}
bp_list$treatment
bp_list[["treatment"]]
bp_list[[1]]
```

```{r}
bp_list$placebo$pre_bp
```

# Task 4

Creation of a status column that holds characters based on length of list

```{r}
bp_list$treatment$status <- character(20)
bp_list$placebo$status <- character(10)
```

For loops and if conditions to assign values based on post values

```{r}
for (i in 1:20){
  if(bp_list$treatment$post_bp[i] <= 120){
    bp_list$treatment$status[i] <- "optimal"
  }
  else if(bp_list$treatment$post_bp[i] <= 130 ){
        bp_list$treatment$status[i] <- "borderline"
  }
  else {
        bp_list$treatment$status[i] <- "high"

  }
}

for (i in 1:10){
  if(bp_list$placebo$post_bp[i] <= 120){
    bp_list$placebo$status[i] <- "optimal"
  }
  else if (bp_list$placebo$post_bp[i] <= 130){
        bp_list$placebo$status[i] <- "borderline"
  }
  else{
        bp_list$placebo$status[i] <- "high"
  }
}

bp_list$treatment
bp_list$placebo
```
