---
title: "hw2"
---

```         
```

# Task 1

This block of code imports the initial BP data, adds identifiers to each data point (Subject_N),

```{r}
#| echo: true
pre_bp = c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)  
post_bp =c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117,101, 119, 130, 122, 106, 106, 124, 102, 117, 113)  
pre_bp
post_bp

subjects=paste("Subject", 1:20, sep = "_")


names(pre_bp)=subjects
names(post_bp)=subjects

pre_bp

bp_difference=pre_bp-post_bp
mean_bp=mean(bp_difference)
mean_bp


decreased = which(bp_difference > 0)
decreased

positive_diff = bp_difference[decreased]
mean(positive_diff)




```

# Task 2

in this task we are simply creating a dataframe for the BP dataset that we have, it just combines all of the data into a singular object that is much easier to manipulate. We also do some filtering of the data, like putting TRUE if the subjects post treatment blood pressure is under 120

```{r}
bp_df = data.frame(
  patient = subjects,
  pre_bp  = pre_bp,
  post_bp = post_bp,
  diff_bp = bp_difference,
  stringsAsFactors = FALSE
)
rownames(bp_df) = NULL
#filters the dataframe for values less than 0, basically the same as using subset
bp_df[bp_df$diff_bp< 0, ]
bp_df$under_120 = bp_df$post_bp < 120
bp_df
knitr::kable(bp_df)

```

# Task 3

Here we are creating a new table with placebo values, formatting it the same format as the blood pressure table. We also place both dataframes in a list, with treatment being the first item in the list and placebo being the second item in the list.

```{r}
#| echo: true
placebo_pre  = c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
placebo_post = c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

placebo_names = paste("Subject", 1:10, sep = "_")

placebo_df = data.frame(
  patient   = placebo_names,
  pre_bp    = placebo_pre,
  post_bp   = placebo_post,
  diff_bp   = placebo_pre - placebo_post,
  stringsAsFactors = FALSE
)
placebo_df$under_120 = placebo_df$post_bp < 120

placebo_df


bp_list <- list(treatment = bp_df, placebo = placebo_df)

bp_list


#accessimg the first element using different types of syntax
bp_list$treatment     
bp_list[["treatment"]] 
bp_list[[1]]          


bp_list$placebo$pre_bp



```

# Task 4

first we are placing the subjects in three different groups based on blood pressure, **optimal** is less than 121, **borderline** is from 121 to 130 inclusive, and **high** is above 130. This is done for both the treatment and the placebo groups.

```{r}

bp_list$treatment$status= character(20)
bp_list$placebo$status= character(10)
#nrow just gets the number of rows, easier than remembering how many rows it has

for (i in 1:nrow(bp_list$treatment)) {
  if (bp_list$treatment$post_bp[i] <= 120) {
    bp_list$treatment$status[i] <- "optimal"
  } else if (bp_list$treatment$post_bp[i] <= 130) {
    bp_list$treatment$status[i] <- "borderline"
  } else {
    bp_list$treatment$status[i] <- "high"
  }
}
bp_list$treatment


#nrow just gets the number of rows, easier than remembering how many rows it has
for (i in 1:nrow(bp_list$placebo)) {
  if (bp_list$placebo$post_bp[i] <= 120) {
    bp_list$placebo$status[i] <- "optimal"
  } else if (bp_list$placebo$post_bp[i] <= 130) {
    bp_list$placebo$status[i] <- "borderline"
  } else {
    bp_list$placebo$status[i] <- "high"
  }
}

bp_list$placebo
```

```{}
```

```{}
```
