---
title: "HW3"
---

1.  The path would be ../MyData.csv

2.  R Projects are good because they allow for the code to run seamlessly no matter who is running the R code, all the individual has to do is have all of the contents of the R project.

3.  Git is a version control system, it tracks the changes of each file in t he repository which allows you to go back to a previous version. Github is a service that hosts git repositories, it stores a copy of a git repository on a server.

4.  A tibble and a data frame are mostly the same, but a tibble returns a tibble when taking a subset, while a dataframe returns a vector. tibbles also only print the first 10 values, while a data frame will default to printing every row and column.

Task 2

```{r}


glass_data = read_csv(
  "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c(
    "id", "refractive_index", "sodium", "magnesium", "aluminum",
    "silicon", "potassium", "calcium", "barium", "iron", "glass_type"
  ),
  show_col_types = FALSE
)
glass_data
```

```{r}
yeast_data = read_table(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
  col_names = c(
    "seq_name", "mcg", "gvh", "alm", "mit",
    "erl", "pox", "vac", "nuc", "class"
  ),
  show_col_types = FALSE
)

yeast_data
```

Task 3

```{r}
download.file(
  "https://www4.stat.ncsu.edu/~online/datasets/white-wine.xlsx",
  destfile = "white-wine.xlsx",
  mode = "wb",
  quiet = TRUE
)

white_wine = read_excel("white-wine.xlsx", sheet = 1)
white_wine


name_key = read_excel("white-wine.xlsx", sheet = 2)


colnames(white_wine) = name_key[[1]]

white_wine$type = "white"

white_wine



red_wine = read_delim(
  "https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
  delim = ";",
  show_col_types = FALSE)


colnames(red_wine) = name_key[[1]]
red_wine$type = "red"

red_wine


wine_data = dplyr::bind_rows(white_wine, red_wine)
wine_data
```

Task 4

```{r}
getIQR=function(x, ...) {
  q = quantile(x, probs = c(0.25, 0.75), ...)
  unname(q[2] - q[1])
}


getIQR(wine_data$fixed_acidity)
getIQR(wine_data$citric_acid)

getCoefVar=function(x, ...){
  sd(x, ...)/ abs(mean(x, ...))
}
getCoefVar(wine_data$fixed_acidity)
getCoefVar(wine_data$citric_acid)


getSummaries=function(x, ...) {
  c(IQR = getIQR(x, ...), CoefVar = getCoefVar(x, ...))}


getSummaries(glass_data$refractive_index)
getSummaries(yeast_data$mcg)

`%cb%` = function(x, y) {
  cbind(x, y)
}


df1=data.frame(unif = runif(10), norm = rnorm(10))
df2=data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1=rnorm(10, mean = 10)

df1 %cb% df2
```
