---
title: "Homework 2 - Grace Haley"
format: html
---
# Task 1

## Question 1

Create vectors for the pre-treatment and post-treatment blood pressure measurements.

```{r}
pre_bp <- c(
  130, 128, 116, 124, 133,
  134, 118, 126, 114, 127,
  141, 138, 128, 140, 137,
  131, 120, 128, 139, 135
)

post_bp <- c(
  114, 98, 113, 99, 107,
  116, 113, 111, 119, 117,
  101, 119, 130, 122, 106,
  106, 124, 102, 117, 113
)
```

## Question 2

Assign subject names to each vector.

```{r}
patient_names <- paste("Subject", 1:20, sep = "_")

names(pre_bp) <- patient_names
names(post_bp) <- patient_names

pre_bp
post_bp
```
## Question 3

Calculate the change in blood pressure for each patient by subtracting the post-treatment measurements from the pre-treatment measurements.

```{r}
diff_bp <- pre_bp - post_bp
diff_bp
```

## Question 4

Calculate the average decrease in blood pressure across all patients.

```{r}
mean(diff_bp)
```

## Question 5

Determine which patients experienced a decrease in blood pressure after treatment using the `which()` function.

```{r}
which(diff_bp > 0)
```

## Question 6

Subset the vector of differences to return only those with a positive change.

```{r}
positive_changes <- diff_bp[diff_bp > 0]

positive_changes
```

## Question 7

Calculate the average decrease in blood pressure for only those patients whose blood pressure decreased.

```{r}
mean(positive_changes)
```

# Task 2

## Question 1

Create a data frame containing the patient IDs, pre-treatment blood pressure, post-treatment blood pressure, and the blood pressure difference.

```{r}
bp_df <- data.frame(
  patient = names(pre_bp),
  pre_bp = as.numeric(pre_bp),
  post_bp = as.numeric(post_bp),
  diff_bp = as.numeric(diff_bp)
)

bp_df
```

## Question 2

Return only the patients whose blood pressure difference is negative.

```{r}
bp_df[bp_df$diff_bp < 0, ]
```

## Question 3

Add a logical column indicating whether the post-treatment blood pressure is below 120.

```{r}
bp_df$post_under_120 <- bp_df$post_bp < 120

bp_df
```

## Question 4

Display the completed data frame as a formatted table.

```{r}
knitr::kable(bp_df)
```

# Task 3

## Question 1

Create a new data frame for the placebo group that includes the patient ID, blood pressure measurements, blood pressure difference, and the `post_under_120` column.

Create vectors for the placebo group

```{r}
pre_bp_placebo <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)

post_bp_placebo <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
```

Create patient names

```{r}
patient_placebo <- paste("Subject", 1:10, sep = "_")
```

Calculate blood pressure difference

```{r}
diff_bp_placebo <- pre_bp_placebo - post_bp_placebo
```

Create the placebo data frame

```{r}
placebo_df <- data.frame(

  patient = patient_placebo,

  pre_bp = pre_bp_placebo,

  post_bp = post_bp_placebo,

  diff_bp = diff_bp_placebo

)
```

Add the post_under_120 column

```{r}
placebo_df$post_under_120 <- placebo_df$post_bp < 120
```

Display the data frame

```{r}
knitr::kable(placebo_df)
```
## Question 2

Create a list containing both data frames

```{r}

bp_list <- list(

  treatment = bp_df,

  placebo = placebo_df

)

bp_list

```

## Question 3

Access the treatment data frame three different ways

```{r}



bp_list[[1]]

bp_list$treatment

bp_list["treatment"]

```

## Question 4
Access the pre-treatment blood pressure column from the placebo data frame

```{r}

bp_list$placebo$pre_bp

```

# Task 4

## Question 1

Create a new `status` column in each data frame stored in the list.

Add an empty status column to the treatment data frame
```{r}
bp_list$treatment$status <- character(nrow(bp_list$treatment))
```
Add an empty status column to the placebo data frame
```{r}
bp_list$placebo$status <- character(nrow(bp_list$placebo))
```
Display the updated list
```{r}
bp_list
```

## Question 2

Assign a blood pressure status to each patient in the treatment group using a for loop.

```{r}
for (i in 1:nrow(bp_list$treatment)) {

  current_bp <- bp_list$treatment$post_bp[i]

  if (current_bp <= 120) {

    bp_list$treatment$status[i] <- "optimal" 

  } else if (current_bp > 120 & current_bp <= 130) {
    
    bp_list$treatment$status[i] <- "borderline"

  } else {

    bp_list$treatment$status[i] <- "high"

  }
}

bp_list$treatment
```

## Question 3

Assign a blood pressure status to each patient in the placebo group using a for loop.

```{r}
for (i in 1:nrow(bp_list$placebo)) {

  current_bp <- bp_list$placebo$post_bp[i]

  if (current_bp <= 120) {

    bp_list$placebo$status[i] <- "optimal" 

  } else if (current_bp > 120 & current_bp <= 130) {
    
    bp_list$placebo$status[i] <- "borderline"

  } else {

    bp_list$placebo$status[i] <- "high"

  }
}

bp_list$placebo
```
---
